Reasoning About a Machine with Local Capabilities

### 2019-02-10 toplas/paper.tex
[TOPLAS] Add reference to attack-defense arms race
... also fix reference.

Compromising software security is often based on attacks that break programming language properties relied upon by software authors such as control-flow correctness, local-state encapsulation, etc.
Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-mitigation arms race) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
The latter techniques essentially emulate protection techniques on existing hardware at the cost of performance, system complexity, and/or security.

\emph{Capability machines} are a type of processors that
remediate these limitations with a better security model at
the hardware level. They are based on old
ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that
 have recently received renewed interest. in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities (a kind
of pointers that carry authority). Memory capabilities carry
the authority to read and/or write to a range of memory
locations. There is also a form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

Unlike commodity processors, capability machines lend themselves well to
enforcing local-state encapsulation. Potentially, they will enable compilation
schemes that enforce this property in an efficient but also 100\% watertight way
(ideally evidenced by a mathematical proof, guaranteeing that we do not end up
in a new attack-mitigation arms race). However, a lot needs to happen before we get
there. For example, it is far from trivial to devise a compilation scheme
adapted to the details of a specific source language's notion of encapsulation
(e.g., private member variables in OO languages often behave quite differently
than private state in ML-like languages). And even if such a scheme were
defined, a formal proof depends on a formalization of the encapsulation provided
by the capability machine at hand.

>>>>

Compromising software security is often based on attacks that break programming language properties relied upon by software authors such as control-flow correctness, local-state encapsulation, etc.
Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-defense arms race \citep{Szekeres_sok_2013}) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
The latter techniques essentially emulate protection techniques on existing hardware at the cost of performance, system complexity, and/or security.

\emph{Capability machines} are a type of processors that remediate these limitations with a better security model at the hardware level.
They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that have recently received renewed interest.
in particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
Capability machines tag every word (in the register file and in memory) to enforce a strict separation between numbers and capabilities (a kind of pointers that carry authority).
Memory capabilities carry the authority to read and/or write to a range of memory locations.
There is also a form of \emph{object capabilities}, which represent the authority to invoke a piece of code without exposing the code's encapsulated private state (e.g., the M-Machine's enter capabilities (described in Section~\ref{sec:capab-mach-with}) or CHERI's sealed code/data pairs).

Unlike commodity processors, capability machines lend themselves well to
enforcing local-state encapsulation. Potentially, they will enable compilation
schemes that enforce this property in an efficient but also 100\% watertight way
(ideally evidenced by a mathematical proof, guaranteeing that we do not end up
in a new attack-defense arms race). However, a lot needs to happen before we get
there. For example, it is far from trivial to devise a compilation scheme
adapted to the details of a specific source language's notion of encapsulation
(e.g., private member variables in OO languages often behave quite differently
than private state in ML-like languages). And even if such a scheme were
defined, a formal proof depends on a formalization of the encapsulation provided
by the capability machine at hand.
